---
title: "use_case_beauty"
author: "Julius Landes"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(mcboost)
#library(fairness)
library(mlr3fairness)
#library(fairml)
#library(fairmodels)
library(data.table)
library(mlr3)
library(mlr3pipelines)
library(mlr3learners)
```
helper function for encoding target variable
```{r}
encode_variable <- function(value) {
  ifelse(value >= 4, 1, 0)
}
```


```{r}
beauty = fread(
  "https://raw.githubusercontent.com/Yorko/mlcourse.ai/master/data/beauty.csv",
  stringsAsFactors = TRUE
)

beauty$looks<-encode_variable(beauty$looks)

beauty$looks <- as.factor(beauty$looks)

beauty$black <- factor(beauty$black, levels = c(0,1), labels = c("non-black","black"))
beauty$female <- factor(beauty$female, levels = c(0,1), labels = c("male","female"))
beauty$union  <- as.factor(beauty$union)
beauty$goodhlth <- as.factor(beauty$goodhlth)
beauty$married <- as.factor(beauty$married )
beauty$service <- as.factor(beauty$service)
tsk_beauty = TaskClassif$new("beauty",beauty, target = "looks")
```
looks von 1-3 -> 0 und 4-5 -> 1 codieren 

Unbalanced groups 
ratio of male female is approx. 2:1 
ratio of blacks / non blacks approx. 1:12 
```{r}
#overview
head(beauty)
str(beauty)
summary(beauty)

#freqs of sensitive vars
table(beauty$black)

table(beauty$female)

```


# splitting ratio 0.8
```{r}
#splitting of data into train and test data
train_tsk_beauty_ind <- partition(tsk_beauty, ratio = 0.8)
#train indices
train_ind <- train_tsk_beauty_ind[[1]]
#test indices
test_ind <- train_tsk_beauty_ind[[2]]
#train data
train_data_beauty<- beauty[train_ind]
#test data
data_test_beauty <- beauty[test_ind]
```

```{r}
train_tsk_beauty = TaskClassif$new("train_data_beauty",train_data_beauty, target = "looks")
```

pipeline for preprocessing
```{r}
pipe_beauty = po("collapsefactors", no_collapse_above_prevalence = 0.0006) %>>%
  po("fixfactors") %>>%
  po("encode") %>>%
  po("imputehist")
prep_task_beauty = pipe_beauty$train(train_tsk_beauty)[[1]]
```

removing sensitive vars
```{r}
prep_task_beauty$set_col_roles(c("black.black","black.non.black","female.female",
                                 "female.male"), remove_from = "feature")
```

fitting random forest
```{r}
l.beauty = lrn("classif.ranger", num.trees = 10L, predict_type = "prob")
l.beauty$train(prep_task_beauty)
```
f0 init predictor
```{r}
init_predictor_beauty = function(data) {
  l.beauty$predict_newdata(data)$prob[, 2]
}
```
extracting data and labels
```{r}
data_beauty = prep_task_beauty$data(cols = prep_task_beauty$feature_names)
labels_beauty = 1 - one_hot(prep_task_beauty$data(cols = prep_task_beauty$target_names)[[1]])
```
multiaccuracy
```{r}
mc_beauty = MCBoost$new(auditor_fitter = "RidgeAuditorFitter", init_predictor = init_predictor_beauty)
mc_beauty$multicalibrate(data_beauty, labels_beauty)
```
multicalibration
```{r}
mcali_beauty = MCBoost$new(
    init_predictor = init_predictor_beauty,
    auditor_fitter = "TreeAuditorFitter",
    num_buckets = 10,
    multiplicative = FALSE
  )

mcali_beauty$multicalibrate(data_beauty, labels_beauty)
```


Evaluation on test data
```{r}

test_tsk_beauty = TaskClassif$new("data_test_beauty", data_test_beauty, target = "looks")
prep_test_beauty = pipe_beauty$predict(test_tsk_beauty)[[1]]
```

```{r}
test_data_beauty = prep_test_beauty$data(cols = prep_test_beauty$feature_names)
test_labels_beauty = 1 - one_hot(prep_test_beauty$data(cols = prep_test_beauty$target_names)[[1]])
```

```{r}
prs_beauty = mc_beauty$predict_probs(test_data_beauty)
```

```{r}
acc_ma_beauty<-mean(round(prs_beauty) == test_labels_beauty)
```

```{r}
acc_rf_beauty<-mean(round(init_predictor_beauty(test_data_beauty)) == test_labels_beauty)
```

```{r}
prs_beauty_cali = mcali_beauty$predict_probs(test_data_beauty)
```

```{r}
acc_mc_beauty<-mean(round(prs_beauty_cali) == test_labels_beauty)
```

```{r}
# Get bias per subgroup for multi-calibrated predictor
data_test_beauty$biasmc = (prs_beauty - test_labels_beauty)
table_mc_beauty<- data_test_beauty[, .(abs(mean(biasmc)), .N), by = .(female,black)]
# Get bias per subgroup for initial predictor
data_test_beauty$biasinit = (init_predictor_beauty(test_data_beauty) - test_labels_beauty)
table_f0_beauty<-data_test_beauty[, .(abs(mean(biasinit)), .N), by = .(female,black)]
```

```{r}
# Get bias per subgroup for multi-calibrated predictor
data_test_beauty$biasmcali = (prs_beauty_cali - test_labels_beauty)
table_mcali_beauty<-data_test_beauty[, .(abs(mean(biasmcali)), .N), by = .(female,black)]
```


```{r}
ae_beauty = mc_beauty$auditor_effect(test_data_beauty)
hist(ae_beauty)
```

3rd iteration used
```{r}
prs_beauty_3 = mc_beauty$predict_probs(test_data_beauty, t = 3L)
```

```{r}
mean(round(prs_beauty_3) == test_labels_beauty)
```
2rd iteration used
```{r}
prs_beauty_2 = mc_beauty$predict_probs(test_data_beauty, t = 2L)
```

```{r}
mean(round(prs_beauty_2) == test_labels_beauty)
```

```{r}
prs_beauty_mcali = mcali_beauty$predict_probs(test_data_beauty)
```

```{r}
mean(round(prs_beauty_mcali) == test_labels_beauty)
```

```{r}
library(formattable)
library(dplyr)
```

```{r}
table_f0_beauty <- table_f0_beauty %>%
  select(black,female,V1,N)
```

data tangling table f0 inital predictor
```{r}
table_f0_beauty[, new_col := paste(black, female)]

table_f0_beauty[, c("black","female") := NULL]

setnames(table_f0_beauty, old = c("new_col","V1"), new = c("subgroups", "classifaction_error"))

setcolorder(table_f0_beauty, neworder = c("subgroups", "classifaction_error", "N"))

table_f0_beauty<- transpose(table_f0_beauty)

setnames(table_f0_beauty, old= c("V1","V2","V3","V4"),
         new = c("non-black female", "non-black male", "black male", "black female"))

table_f0_beauty<-table_f0_beauty[-1]
```

data tangling table mc beauty
```{r}
table_mc_beauty[, new_col := paste(black, female)]

table_mc_beauty[, c("black","female") := NULL]

setnames(table_mc_beauty, old = c("new_col","V1"), new = c("subgroups", "classifaction_error"))

setcolorder(table_mc_beauty, neworder = c("subgroups", "classifaction_error", "N"))

table_mc_beauty<- transpose(table_mc_beauty)

setnames(table_mc_beauty, old= c("V1","V2","V3","V4"),
         new = c("non-black female", "non-black male", "black male", "black female"))

table_mc_beauty<-table_mc_beauty[-1]

table_beauty <- rbind(table_f0_beauty, table_mc_beauty)

table_beauty <- table_beauty[-2]
```

```{r}
table_beauty_round<-table_beauty[1:2]
table_beauty_groups <- table_beauty[3]
table_beauty_round<-table_beauty_round[, lapply(.SD, function(x) round(as.numeric(x), 3))]
```

```{r}
table_beauty<-rbind(table_beauty_groups, table_beauty_round)

new_row_names <- c("group size", "RF", "MA")
rownames(table_beauty) <- new_row_names
```

```{r}
beauty_table <- formattable(table_beauty)
```

# splitting ratio 0.5
```{r}
#splitting of data into train and test data
train_tsk_beauty_ind <- partition(tsk_beauty, ratio = 0.5)
#train indices
train_ind <- train_tsk_beauty_ind[[1]]
#test indices
test_ind <- train_tsk_beauty_ind[[2]]
#train data
train_data_beauty<- beauty[train_ind]
#test data
data_test_beauty <- beauty[test_ind]
```

```{r}
train_tsk_beauty = TaskClassif$new("train_data_beauty",train_data_beauty, target = "looks")
```

pipeline for preprocessing
```{r}
pipe_beauty = po("collapsefactors", no_collapse_above_prevalence = 0.0006) %>>%
  po("fixfactors") %>>%
  po("encode") %>>%
  po("imputehist")
prep_task_beauty = pipe_beauty$train(train_tsk_beauty)[[1]]
```

removing sensitive vars
```{r}
prep_task_beauty$set_col_roles(c("black.black","black.non.black","female.female",
                                 "female.male"), remove_from = "feature")
```

fitting random forest
```{r}
l.beauty = lrn("classif.ranger", num.trees = 10L, predict_type = "prob")
l.beauty$train(prep_task_beauty)
```
f0 init predictor
```{r}
init_predictor_beauty = function(data) {
  l.beauty$predict_newdata(data)$prob[, 2]
}
```
extracting data and labels
```{r}
data_beauty = prep_task_beauty$data(cols = prep_task_beauty$feature_names)
labels_beauty = 1 - one_hot(prep_task_beauty$data(cols = prep_task_beauty$target_names)[[1]])
```
multiaccuracy
```{r}
mc_beauty = MCBoost$new(auditor_fitter = "RidgeAuditorFitter", init_predictor = init_predictor_beauty)
mc_beauty$multicalibrate(data_beauty, labels_beauty)
```
multicalibration
```{r}
mcali_beauty = MCBoost$new(
    init_predictor = init_predictor_beauty,
    auditor_fitter = "TreeAuditorFitter",
    num_buckets = 10,
    multiplicative = FALSE
  )
mcali_beauty$multicalibrate(data_beauty, labels_beauty)
```


Evaluation on test data
```{r}

test_tsk_beauty = TaskClassif$new("data_test_beauty", data_test_beauty, target = "looks")
prep_test_beauty = pipe_beauty$predict(test_tsk_beauty)[[1]]
```

```{r}
test_data_beauty = prep_test_beauty$data(cols = prep_test_beauty$feature_names)
test_labels_beauty = 1 - one_hot(prep_test_beauty$data(cols = prep_test_beauty$target_names)[[1]])
```

```{r}
prs_beauty = mc_beauty$predict_probs(test_data_beauty)
```

```{r}
acc_beauty_ma<-mean(round(prs_beauty) == test_labels_beauty)
```

```{r}
prs_beauty_cali = mcali_beauty$predict_probs(test_data_beauty)
```

```{r}
acc_beauty_mc<- mean(round(prs_beauty_cali) == test_labels_beauty)
```
accuracy of f0
```{r}
acc_f0_beauty<-mean(round(init_predictor_beauty(test_data_beauty)) == test_labels_beauty)
```

```{r}
# Get bias per subgroup for multi-calibrated predictor
data_test_beauty$biasmc = (prs_beauty - test_labels_beauty)
table_mc_beauty<- data_test_beauty[, .(abs(mean(biasmc)), .N), by = .(female,black)]
# Get bias per subgroup for initial predictor
data_test_beauty$biasinit = (init_predictor_beauty(test_data_beauty) - test_labels_beauty)
table_f0_beauty<-data_test_beauty[, .(abs(mean(biasinit)), .N), by = .(female,black)]
```

```{r}
ae_beauty = mc_beauty$auditor_effect(test_data_beauty)
hist_08<-hist(ae_beauty)
```

```{r}
table_f0_beauty <- table_f0_beauty %>%
  select(black,female,V1,N)
```

data tangling table f0 inital predictor
```{r}
table_f0_beauty[, new_col := paste(black, female)]

table_f0_beauty[, c("black","female") := NULL]

setnames(table_f0_beauty, old = c("new_col","V1"), new = c("subgroups", "classifaction_error"))

setcolorder(table_f0_beauty, neworder = c("subgroups", "classifaction_error", "N"))

table_f0_beauty<- transpose(table_f0_beauty)

setnames(table_f0_beauty, old= c("V1","V2","V3","V4"),
         new = c("non-black female", "non-black male", "black male", "black female"))

table_f0_beauty<-table_f0_beauty[-1]
```

data tangling table mc beauty
```{r}
table_mc_beauty[, new_col := paste(black, female)]

table_mc_beauty[, c("black","female") := NULL]

setnames(table_mc_beauty, old = c("new_col","V1"), new = c("subgroups", "classifaction_error"))

setcolorder(table_mc_beauty, neworder = c("subgroups", "classifaction_error", "N"))

table_mc_beauty<- transpose(table_mc_beauty)

setnames(table_mc_beauty, old= c("V1","V2","V3","V4"),
         new = c("non-black female", "non-black male", "black male", "black female"))

table_mc_beauty<-table_mc_beauty[-1]

table_beauty <- rbind(table_f0_beauty, table_mc_beauty)

table_beauty <- table_beauty[-2]
```

```{r}
table_beauty_round<-table_beauty[1:2]
table_beauty_groups <- table_beauty[3]
table_beauty_round<-table_beauty_round[, lapply(.SD, function(x) round(as.numeric(x), 3))]
```

```{r}
table_beauty<-rbind(table_beauty_groups, table_beauty_round)

table_mcali_beauty_values <- table_mcali_beauty$V1

table_mcali_beauty_values_r <-round(table_mcali_beauty_values, digits = 3)

table_mcali_beauty_values_r2 <- t(table_mcali_beauty_values_r)
table_mcali_beauty_values_r3 <- as.data.frame(table_mcali_beauty_values_r2)

setnames(table_mcali_beauty_values_r3, old= c("V1","V2","V3","V4"),
         new = c("non-black female", "non-black male", "black male", "black female"))
table_beauty_final <- rbind(table_beauty, table_mcali_beauty_values_r3)
table_beauty_final$Overall_Accuracy <- c(NA,1-round(acc_f0_beauty, digits = 3), 1-round(acc_beauty_ma, digits = 3), 1-round(acc_beauty_mc, digits = 3))
table_beauty_final[is.na(table_beauty_final)] <- ""

new_row_names <- c("group size", "RF", "MA","MC")
rownames(table_beauty_final) <- new_row_names

col_names <- c("non-black female", "non-black male", "black male", "black female", "Overall Classification Error")
colnames(table_beauty_final) <- col_names
```

```{r}
beauty_table_05 <- formattable(table_beauty_final)
```
